Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-01-15 doc/article.tex
finish draft of article

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb,amsmath}
\usepackage{natbib}
\usepackage{grffile}
\usepackage{textcase}
\usepackage{hyperref}
%SetFonts

@@

The projected precisions for LSST-discovered SNe~Ia have a number of interesting features. 
Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that 
there is little benefit in going beyond
$z=0.3$.
At $z<0.2$ the constraining power on  $f\sigma_8$ 
comes primarily from peculiar velocities
but by $z=0.3$ RSD is more important.
It turns out that for $z>0.1$, the volume is sufficiently large that the LSST sample is not yet sample variance limited.

\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\subsection{Survey Parameters that Determine Scientific Performance}

The expected peculiar velocity signal  is described by its power spectrum $P_{vv}$.
The range $k$'s  used in an analysis depends on the depth of the survey $k_{min} \approx (2 z_{max})^{-1}$, and the smallest
scales where the theoretical predictions are reliable, say $k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$.
The top panel of Figure~\ref{detC:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
at $z=0$ assuming General Relativity and linear theory.  The panel shows available $k$ ranges for several surveys with different
maximum redshift: the $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes contribute the bulk of the peculiar velocity signal.
The sensitivity to gravity primarily resides in the normalization of the power spectrum, as  $P_{vv} \propto (f\sigma_8)^2$.
Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$  in Figure~\ref{power:fig}.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{src/power.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{ 
\label{power:fig}}
\end{figure}

\citet{2017ApJ...847..128H} project how complementary galaxy and peculiar velocity surveys in a common volume can measure $f\sigma_8$. 
They perform their analysis in Fourier space, 
where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C$ is
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right]
\label{fisher:eqn}
\end{equation}
and the covariance is 
\begin{equation}
C(k,\mu)  =
  \begin{bmatrix}
   P_{gg}(k,\mu) + \frac{1}{n_g} &
   P_{vg}(k,\mu)  \\
   P_{vg}(k,\mu)  &
  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
   \end{bmatrix}.
\label{cov:eq}
\end{equation}
The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the bias parameter.
The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the probe velocity dispersion
is $\sigma$. 
 The Fisher matrix in  Eq.~\ref{fisher:eqn} is evaluated by \citet{2017MNRAS.464.2517H}  to generate the $f \sigma_8$ uncertainties in Table~\ref{tab:howlett}.

The integrand
\begin{equation}
I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
\frac{\partial C}{\partial  (f\sigma_8)} \right]
\end{equation}
in Eq.~\ref{fisher:eqn}  represents how much a differential Fourier
volume element with wavenumber $k$ and line of sight $\mu$ contributes to $F_{f\sigma_{8}\,f\sigma_{8}}$;
it is therefore  useful in evaluating the design of peculiar velocity surveys. 
We evaluate these surveys relative to the $I_{max}$ of a sample-variance limited
peculiar velocity survey with\ $\sigma/n =0$.
In the following we consider as given a  TAIPAN-like
galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ that overlaps with the peculiar velocity surveys.

First, we show the benefit of having the  galaxy and velocity surveys in a common volume by considering the degradation that occurs when 
they are treated as being
independent.  For this exercise the velocity survey is taken to be sample-variance limited.
The resulting $I/I_{max}$  is shown as the black dashed curves in the bottom two panels of Figure~\ref{power:fig},
and is found to lie between $0.1-0.2$.
Taking advantage of the fact that the same dark-matter fluctuations produce the galaxies and velocities 
in the common volume decreases
the variance in $f\sigma_8$ by factors of 5 to 10.

Realistic surveys have shot noise. We consider two- and ten-year
supernova densities $n$  based on SN~Ia rates 
with 0.65 sample efficiency, and velocity dispersions $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$
assuming a supernova magnitude dispersion of $\sigma_M=0.08$ and inserting values of $z=0.05,0.1, 0.2$.
The $I/I_{max}$ of these survey scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ meaning that they are sample-noise limited.
Individual $k>0.01$ modes have  $I/I_{max}<1$ showing that shot noise
is significant.  Recall that collectively,  larger $k$ modes are the important contributors to the peculiar-velocity signal;
basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$..
The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).

The survey parameters that  enter the Fisher matrix are the volume $V$, which
in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a fixed parameter, which only increases with survey duration.
Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.

\begin{itemize}
\item Solid Angle $\Omega$:
The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole, could significantly increase
the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
\item Redshift Depth $z_{max}$:
For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
limited.  We have run simulations of LSST SN surveys
and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
%
%\begin{figure}
%\centering
%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
%realizations made in simulating the surveys.
%For several limiting redshift depths, the STON is plotted as a
%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
%\label{notshot:fig}}
%\end{figure}

The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers, is higher for lower
$z_{max}$.  This is shown in the
right plot of   Figure~\ref{notshot:fig}.

A low-redshift tracer is more valuable than one at higher redshift.

\item Intrinsic Magnitude Dispersion $\sigma_M$:
A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$,
as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
limited so the $\sigma^2_M/n$-term
in the Fisher matrix is important.
Survey follow-up strategy,
which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration.

The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
a pure SN~Ia sample obtained through
spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning, particularly in extending the redshift range.
\end{itemize}

@@

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence  
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
reach a targeted $\sigma_M$.
\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.

@@

@@

>>>>

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb,amsmath}
\usepackage{natbib,verbatim}
\usepackage{grffile}
\usepackage{textcase}
\usepackage{hyperref}
%SetFonts

@@

The projected precisions for LSST-discovered SNe~Ia have a number of interesting features. 
Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that 
there is little benefit in going beyond
$z=0.3$.
At $z<0.2$ the constraining power on  $f\sigma_8$ 
comes primarily from peculiar velocities
but by $z=0.3$ RSD is more important.

\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\subsection{Fisher Matrix of Velocity-Galaxy Surveys}

The  peculiar velocity signal  can be described by its power spectrum $P_{vv}$.
The sensitivity to gravity primarily resides in the normalization of the power spectrum, as  $P_{vv} \propto (f\sigma_8)^2$.
Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
though this detail is not considered here.
The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
at $z=0$ expected from General Relativity and linear theory. Available $k$ ranges for surveys with different
maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
scale where  theory predictions are reliable.
Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{src/power.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Top panel: Volume-weighted velocity power spectrum $k^3P_{vv}$ at $z=0$ for $\mu=1, 0,5$ as predicted for General Relativity in the linear regime.
Lower two panels: $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
redshift depths $z_{max}$.
\label{power:fig}}
\end{figure}

\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
 obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
Their analysis is performed in Fourier space, 
where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right]
\label{fisher:eqn}
\end{equation}
 and covariance
\begin{equation}
C(k,\mu)  =
  \begin{bmatrix}
   P_{gg}(k,\mu) + \frac{1}{n_g} &
   P_{vg}(k,\mu)  \\
   P_{vg}(k,\mu)  &
  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
   \end{bmatrix}.
\label{cov:eq}
\end{equation}
The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
is $\sigma$. 
The integrand in the Fisher-element calculation
\begin{equation}
I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
\frac{\partial C}{\partial  (f\sigma_8)} \right]
\end{equation}
represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
differential Fourier
volume element.

\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
The performance of a peculiar velocity survey depends on its overlap with galaxy surveys, solid-angle coverage, redshift depth, and intrinsic magnitude dispersion.
In the following, we consider as given a  TAIPAN-like
galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ that overlaps with the peculiar velocity surveys.
While the expected variance in $f\sigma_8$ is $\left(F^{-1}\right)_{{f\sigma_{8}\,f\sigma_{8}}}$, we use $I$ as a proxy to help understand
how survey densities and shot noise affect science performance.
Surveys are compared relative to the $I_{max}$ of an idealized sample-variance limited peculiar-velocity survey with\ $\sigma/n =0$.

First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
Consider two- and ten-year
SN~Ia densities $n$  based on SN~Ia rates 
with 0.65 efficiency, and velocity dispersions $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$
assuming a supernova magnitude dispersion of $\sigma_M=0.08$ and inserting values of $z=0.05,0.1, 0.2$.
The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
%The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).

{\bf Solid Angle $\Omega$:}
$F_{f\sigma_{8}\,f\sigma_{8}}$ is  proportional to volume and hence the survey solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole could significantly increase
the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.

%s
% and so helps illustrate 
%how how the density, $n$, and intrinsic SN magnitude dispersion of peculiar velocity probes, as
%$\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, affect survey precision.

{\bf Common-Volume Galaxy  Survey:}
Galaxy and velocity surveys in different volumes do not have the off-diagonal cross-correlation terms in Eq.~\ref{cov:eq}.
The degradation in probative power   $I/I_{max}$, when sample-variance limited,
is shown as the black dashed curves in the bottom two panels of Figure~\ref{power:fig}.
The curves lie between $0.1-0.2$, showing
that accounting for the common dark-matter fluctuations that give rise to galaxy densities and peculiar velocities
reduces the variance in $f\sigma_8$ by factors of 5 to 10.

{\bf Redshift Depth $z_{max}$:}
Increasing the survey redshift depth increases the $f \sigma_8$ precision, but when resource-limited is the expense of additional depth worth neglecting lower redshift
SNe? The answer is found in  Table~\ref{tab:howlett}, which shows that after 10 years of LSST the lowest redshifts have the
strongest probative power and are sensitive more to peculiar velocities rather than galaxy overdensities.  Despite
the differential volume increase with redshift depth (Eq,~\ref{fisher:eqn}), the degradation in intrinsic peculiar velocity $\propto z/(1+z)$
and being shot-noise-limited  makes a low-redshift tracer more valuable than one at higher redshift.

{\bf Intrinsic Magnitude Dispersion $\sigma_M$}:
Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
%limited so the $\sigma^2_M/n$-term
%in the Fisher matrix is important.
Survey follow-up strategy,
which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration.

%\subsection{Survey Parameters that Determine Scientific Performance}

%The survey parameters that  enter the Fisher matrix are the volume $V$, which
%in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
%magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.

%
%
%\begin{itemize}
%\item Solid Angle $\Omega$:
%The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
%LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
%the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
%coverage of the southern equatorial pole, could significantly increase
%the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
%\item Redshift Depth $z_{max}$:
%Increasing the survey redshift depth increases the $f \sigma_8$ precision, but when resource-limited is it worth doing so at the expense of neglecting lower redshift
%SNe? This is answered in  Table~\ref{tab:howlett}, which shows that after 10 years of LSST the lowest redshifts have the
%stronger probative power and benefit most from peculiar velocities.  Although
%a differential increase in redshift depth increases the volume in Eq,~\ref{fisher:eqn}, the degradation in  intrinsic peculiar velocity $\propto z/(1+z)$
%and the fact that shot noise is still important leads to the conclusion that a low-redshift tracer is more valuable than one at higher redshift.
%
%%For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
%%limited.  We have run simulations of LSST SN surveys
%%and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
%%of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
%%It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
%%
%%\begin{figure}
%%\centering
%%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
%%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
%%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
%%realizations made in simulating the surveys.
%%For several limiting redshift depths, the STON is plotted as a
%%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
%%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
%%\label{notshot:fig}}
%%\end{figure}
%%
%%The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers, is higher for lower
%%$z_{max}$.  This is shown in the
%%right plot of   Figure~\ref{notshot:fig}.
%%
%
%
%\item Intrinsic Magnitude Dispersion $\sigma_M$:
%Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
%A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
%%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
%%limited so the $\sigma^2_M/n$-term
%%in the Fisher matrix is important.
%Survey follow-up strategy,
%which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
%This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
% in number density, or equivalently in survey duration. 
%\end{itemize}

@@

%Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a fixed parameter, which only increases with survey duration.
%Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
a pure SN~Ia sample obtained through
spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning, particularly in extending the redshift range.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
The nominal cadence  
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
reach a targeted $\sigma_M$.
\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.

@@

\begin{comment}

@@

\end{comment}
